---
title: "Basic zoon usage"
author: "Tim Lucas"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    theme: peaksea
    highlight: zenburn
  pdf_document:
    theme: peaksea
    highlight: zenburn
---

<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Basic zoon usage}
-->


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra='style="display:block; margin: auto"', fig.align="center")
```

An Introduction to the **zoon** package
=======================================


**Zoon** is a package to aid reproducibility and between-model comparisons in species distribution modelling. Each step in an analysis is a 'module'. These modules will include: 
+ Data collection of **occurrence** and environmental **covariate** data from online databases. 
+ **Process** steps such as removal of spatial autocorrelation in the data or generation of background pseudoabsences.
+ The fitting of **models**.
+ Model **output** including diagnostics, reports and vizualisation.



Basic usage
----------------------------

First install from github
```{r methods, eval = FALSE}
library(devtools)
install_github('zoonproject/zoon')
```

and load

```{r load}
library(zoon)
```


