<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{An Introduction to the zoon package}
-->

An Introduction to the **zoon** package
=======================================


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(out.extra='style="display:block; margin: auto"', fig.align="center")
```


Introduction
------------


**Zoon** is a package to aid reproducibility and between model comparisons in species distribution modelling. The aim is for each step in an analysis to be a 'module'. These modules will include: 
+ Data collection of occurrence and environmental covariate data from online databases. 
+ Preprocessing steps such as removal of spatial autocorrelation in the data or generation of background pseudoabsences.
+ The fitting of models.
+ Model diagnostics.
+ Model output and vizualisation.

We aim for the creation of these modules to be as simple as possible.


Basic usage
----------------------------

First install from github
```{r methods}
#library(devtools)
#install_github('zoonproject/zoon')

library(zoon)
```


Then run two workflows. Both workflows use the module 'AnophelesPlumbeus' to get \emph{Anopheles plumbeus} records from GBIF. 'AirNCEP' then collects air temperature records from NCEP. The module 'OneHundredBackground' uses 100 random bakground points as our pseudoabsence data. 

The two workflows run different SDMs. The first runs a logistic regression while the second runs a random forest. Finally, we use the module 'SameTimePlaceMap' to predict occurence across the current extent we already defined and for the present time.
```{r main analysis, eval=TRUE, cache=FALSE}
# run a workflow, using the logistic regression model
#ans1 <- workflow(occurMod = 'UKAnophelesPlumbeus',
 #                covarMod = 'UKAir',
 #                procMod = 'OneHundredBackground',
 #                modelMod = 'LogisticRegression',
 #                outMod = 'SameTimePlaceMap')

# switch the model to a RandomForest
#ans2 <- workflow(occurMod = 'UKAnophelesPlumbeus',
#                 covarMod = 'UKAir',
#                 procMod = 'OneHundredBackground',
#                 modelMod = 'RandomForest',
#                 outMod = 'SameTimePlaceMap')



```
Then we can have a look at our results.


```{r output, eval=TRUE, cache=TRUE}
# look at the contents of these lists
str(ans1, 1)
str(ans2, 1)

# plot the resulting maps
par(mfrow = c(1, 2))

plot(ans1$output[[1]],
     zlim = c(0,1),
     main = 'LR')

points(ans1$occurrence.output[[1]][, 1:2],
       pch = 16,
       cex = 0.3)

plot(ans2$output[[1]],
     zlim = c(0,1),
     main = 'RF')

points(ans2$occurrence.output[[1]][, 1:2],
       pch = 16,
       cex = 0.3)
```

Some points to note about reproducibility. The whole output object should contain all the information needed to reproduce the analysis. So for example, the data is saved. Therefore, even if GBIF gets new data, the exact data used in this analysis is stored. 

```{r storedData}
head(ans1$occurrence.output)
```



