---
title: "Building modules"
author: "Tim Lucas"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    theme: peaksea
    highlight: zenburn
  pdf_document:
    theme: peaksea
    highlight: zenburn
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Building modules}
-->


# Building a module

The process of making a module is essentially 
1. Write an R function
2. Run `BuildModule` with the function and metadata

and optionally
3. Upload to the zoon modules repository

## An example

Here is a simple function that will become our module. It is a model module that uses general additive models.

```{r newMod, eval = FALSE}
NewModule <- function(df){

  zoon:::GetPackage("gam")
  
  covs <- as.data.frame(df[, 6:ncol(df)])
  names(covs) <- names(df)[6:ncol(df)]
  m <- gam::gam(formula = df$value ~ .,
         data = covs,
         family = binomial)

}
```
We then run `BuildModule` adding fairly extensive meta data and directing BuildModule to save the file in the working directory '.'. As this module has no parameters other than `df` which is not user specified, set `paras` to list().

```{r BuildMod, eval = FALSE}

BuildModule(NewModule, type = 'model', dir = '.', title = 'GAM sdm model', 
  description = 'This is my mega cool new model. It does GAMs using the gam package.',
  paras = NULL, author = 'Z. Oon', email = 'zoon@zoon.com')

```

This is now a runeable module.

```{r newmodworkflow, eval = FALSE}
rm(NewModule)
LoadModule('NewModule.R')
work1 <- workflow(occurrence = UKAnophelesPlumbeus,
                  covariate = UKAir,
                  process  = OneHundredBackground,
                  model = NewModule,
                  output   = PrintMap)
```

Once we're happy with the module, we will hopefully upload it to the zoon repository.


# Module IO definitions for module developers

The input arguments and return values of modules are strict. However, any module type can have additional input arguments but these must be named. A lot of the data frames include '+ covariates'. This indicates that there the number of covariate columns is flexible.

### Occurrence
Out: dataframe with column names: longitude, latitude, value, type, fold

### Covariate
out: raster layer or raster stack

### Process
in: list( df with values, type, fold, longitude, latitude + covariates,  
          covariate rasterstack/layer
        )

out:  list(df with values, type, fold, longitude, latitude + covariates, 
           covariate rasterstack/layer
          )

### Model
in: dataframe from process
out: model object 
(Object with a predict method. If not, please define a method within the module. See BiomodModel for an example.)

### Output
in: list: list$model (model object from above) and list$data is a dataframe from process.output + predictions. Rasterlayer 
out: Anything

# Pictoral description of inputs and outputs
![OccurrenceModule](occurrenceInOut.svg)
![CovariateModule](covariateInOut.svg)
![ProcessModule](processInOut.svg)
![ModelModule](modelInOut.svg)
![OuputModule](outputInOut.svg)




