---
title: "Building modules"
author: "Tim Lucas & Tom August"
date: "`r Sys.Date()`"
output: html_vignette
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Building modules}
-->


# Building a module

The process of making a module is essentially

1. Write an R function
2. Run `BuildModule` with the function and metadata
3. *Optional* -- Upload to the zoon modules repository

## An example

Here is a simple function that will become our module. It is a model module that uses general additive models. We will work through it one element at a time

First we start our function by declaring all the parameters we need, including all the defaults

```{r, eval = FALSE}
NewModule <- function(.df){
```

Since this is a model module the only default is `.df`. To find out more about defaults see the section [Module IO definitions for module developers](#tag1).

Next we specify the packages our function needs. These should be specified by using GetPackage function in the zoon package. This function will load the package if the user of your module already has it or will install it from CRAN if they don't. For this reason make sure your package only uses packages that are on CRAN.

```{r, eval = FALSE}
# Specify the packages we need using the function
# GetPackage
zoon:::GetPackage("gam")
```  

Next we can add the code that does our modelling, here we create a simple GAM (Generalised Additive Model) using the package [gam](https://cran.r-project.org/web/packages/gam/index.html)

```{r, eval = FALSE}
# Create a data.frame of covariate data
covs <- as.data.frame(.df[, 6:ncol(.df)])
names(covs) <- names(.df)[6:ncol(.df)]

# Run our gam model
m <- gam::gam(formula = .df$value ~ .,
              data = covs,
              family = binomial)
```

The final stage of building a model module is to create a ZoonModel object. This is important as it ensures that all model module output are the same and specifically that zoon can predict from them easily.

We build a ZoonModel object by using the function `ZoonModel`. This takes three parameters

1. *model* - Your model object
2. *code* - A section of code that will use `model` [your model] and `newdata` [a new set of covariate data], to return a vector of predicted values, one for each row of `newdata`
3. *packages* - A vector of characters naming the packages needed to run *code*

```{r, eval = FALSE}
# Create a ZoonModel object to return.
# this includes our model, predict method
# and the packages we need.
ZoonModel(model = m,
          code = {
          
          # create empty vector of predictions
          p <- rep(NA, nrow(newdata))
          
          # omit NAs in new data
          newdata_clean <- na.omit(newdata)
          
          # get NA indices
          na_idx <- attr(newdata_clean, 'na.action')
          
          # if there are no NAs then the index should 
          # include all rows, else it should name the 
          # rows to ignore
          if (is.null(na_idx)){
            na_idx <- 1:nrow(newdata)
          } else {
            idx <- -na_idx
          }
          
          # Use the predict function in gam to predict
          # our new values
          p[idx] <- gam::predict.gam(model,
                                     newdata_clean,
                                     type = 'response')
          return (p)
        },
        packages = 'gam')
```

With all these elements in place we now have our module complete. All together it looks like this.

```{r, eval = FALSE}
NewModule <- function(.df){

  # Specify the packages we need using the function
  # GetPackage
  zoon:::GetPackage("gam")
  
  # Create a data.frame of covariate data
  covs <- as.data.frame(.df[, 6:ncol(.df)])
  names(covs) <- names(.df)[6:ncol(.df)]
  
  # Run our gam model
  m <- gam::gam(formula = .df$value ~ .,
         data = covs,
         family = binomial)
  
  # Create a ZoonModel object to return.
  # this includes our model, predict method
  # and the packages we need.
  ZoonModel(model = m,
            code = {
            
            # create empty vector of predictions
            p <- rep(NA, nrow(newdata))
            
            # omit NAs in new data
            newdata_clean <- na.omit(newdata)
            
            # get their indices
            na_idx <- attr(newdata_clean, 'na.action')
            
            # if there are no NAs then the index should 
            # include all rows, else it should name the 
            # rows to ignore
            if (is.null(na_idx)){
              na_idx <- 1:nrow(newdata)
            } else {
              idx <- -na_idx
            }
            
            # Use the predict function in gam to predict
            # our new values
            p[idx] <- gam::predict.gam(model,
                                       newdata_clean,
                                       type = 'response')
            return (p)
          },
          packages = 'gam')
  
}
```

We then run `BuildModule` adding fairly extensive meta data and directing BuildModule to save the file in the working directory '.'. As this module has no parameters other than `.df` which is not user specified, set `paras` to list(). Default parameters, like `.df`, all start with with a `.` and will be written into the module documentation automatically.

```{r BuildMod, eval = FALSE}
BuildModule(object = NewModule,
            type = 'model',
            dir = '.',
            title = 'GAM sdm model',
            description = 'This is my mega cool new model.',
            details = 'This module performs GAMs (Generalised Additive Models) using the \\code{gam} function from the package \\code{gam}.',
            paras = NULL,
            author = 'Z. Oon',
            email = 'zoon@zoon.com')
```

This is now a run-able module.

```{r newmodworkflow, eval = FALSE}
rm(NewModule)
LoadModule('NewModule.R')
work1 <- workflow(occurrence = UKAnophelesPlumbeus,
                  covariate = UKAir,
                  process  = OneHundredBackground,
                  model = NewModule,
                  output   = PrintMap)
```

Once we're happy with the module, we will hopefully upload it to the zoon repository. The repository is currently under development. Visit [the development pages](https://zoonproject.wordpress.com/) for more information.


# <a name="tag1"></a>Module IO definitions for module developers

The input arguments and return values of modules are strict. However, any module type can have additional input arguments but these must be named. A lot of the data frames include '+ covariates'. This indicates that the number of covariate columns is flexible.

### Occurrence
In: No default inputs

Out: data.frame with column names: longitude, latitude, value, type, fold

### Covariate
In: No default inputs

Out: raster layer or raster stack

### Process
In: list called **.data** with 2 elements

- *df* - A data.frame with columns: values, type, fold, longitude, latitude + covariates
- *ras* - A covariate rasterstack/layer

Out:  list with 2 elements

- *df* - A data.frame with columns: values, type, fold, longitude, latitude + covariates
- *ras* - A covariate rasterstack/layer

### Model
In: data.frame from process called **.df**

Out: A ZoonModel object (see the example above)

### Output
In: list called **.model** with 2 elements

- *model* - A ZoonModel object from a model module
- *data* - A data.frame from a process module with the added column `predictions`

Also a Rasterlayer called **.ras** from the covariate module

Out: Anything

# Pictoral description of inputs and outputs
![OccurrenceModule](occurrenceInOut.svg)
![CovariateModule](covariateInOut.svg)
![ProcessModule](processInOut.svg)
![ModelModule](modelInOut.svg)
![OuputModule](outputInOut.svg)




